---
title: "Practical 3"
output: 
  html_document:
    css: style.css
---

```{r setup practical 3, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(here)
library(RVAideMemoire)
library(tidyverse)
library(kableExtra)
library(gridExtra)
library(extRemes)

```

# Practical 3

In this third practical, we expand the retail sales dataset of Practical 2 to include multiple products. Data for Product 1 (from Practical 2) and an additional 9 products are given in two forms: $sales10.csv$ and $sales10adjusted.csv$, with the same format. 
As the name indicates, sales volumes in the latter file have been adjusted for trend and seasonality *(removing growth and cyclical patterns).*

```{r }
#data

sales10 <- read.csv(here::here("sales_10.csv"))
sales10adj <- read.csv(here::here("sales_10_adjusted.csv"))

```

## Exercise A

> Explain the challenges involved when analysing the risk of stock-out across several products. What is the danger of considering each series independently?

The aplication of an independently analysis (EVT) could cause a loss of information and therefore a lack of preparation in face of a risky event. For example, the analysis of stock out from 3 products independently are not consider risky, nevertheless, the stock out of the 3 together could be consider as a very high risky event due to the high probability of loosing a customer. 

**MEVT has as a main objective to find the probabilities that extreme events happen together.**

The main problem would be given by the fact that the different products may not be independent even if we consider them to be, it can be seen if one of them has an extreme value in the demand and so does another one, every time at the same date. 
This can happen as the products may be related, as for example could be bread and butter, meaning that if there is an increase in the demand of one product, so there will be in the demand for the other one. 
Another possibility is the fact that, especially for extreme values, so an abrupt change in the demand, could be determined by external factors, such as a natural disaster, that would make the demand for all sort of products increase from one day to the other. 
Indeed, it could be a problem, as by analysing the stock-outs, which are extreme values, independently one from the other, we could miss some important relationship that are between them and not being able to assess the risk in the best way possible.

## Exercise B

> Propose one graphical and one numerical method of detecting dependence of extreme values of the demand across several products. Apply your chosen methods to the adjusted sales data and identify groups of related products (if any exist).

| n° | Analysis     |  Method                                                           | 
|:--:|:-------------|:------------------------------------------------------------------|   
| 1  | Graphical    | We will apply two methods: <br /> 1) A plot with **all the products distributions** over time, in which we will set a threshold as an arbitrary value, in order to enable us the study the ones which are above it and see if there is repeated pattern (for example having always two products that are over the threshold in two close dates) <br /> 2) A **chi-plot** among all the different couples of products, so that we can study the distribution of the tails of the products' demands. |
| 2  | Numerical   | We will **compute the chibar values**, which will confirm us the tail dependences among all the different couples of products. |
| 3  | Aggregation | In this step we will focus in the grouping of products, in order to do so, we will use the following 2 method: <br /> 1) We will calculate the mean and variance for each product and display them on a graph, to see if they have any other product close to them <br /> 2) Then, we will compute the Kendall-Taus distances among the products and see the clustering of the products based on it.|


### Graphical method {.tabset .tabset-fade .tabset-pills}

First, let's create a plot with all the products' demands distibutions. 

#### All the products distribution {-}

```{r, echo=FALSE, fig.height=8, message=FALSE, warning=FALSE}

#plot with all products' demand
plot(sales10adj$Product_1[1:1500], type = "l", 
                        col = "blue", main = "Daily Sales",xlab = "Day", 
                        ylab = "Daily sales (unit)") +
                  lines(sales10adj$Product_2[1:1500],  type = "l", col = "green") +
                  lines(sales10adj$Product_3[1:1500], type = "l", col = "purple") +
                  lines(sales10adj$Product_4[1:1500], type = "l", col = "red") +
                  lines(sales10adj$Product_5[1:1500], type = "l", col = "grey") +
                  lines(sales10adj$Product_6[1:1500], type = "l", col = "brown") +
                  lines(sales10adj$Product_7[1:1500], type = "l", col = "orange") +
                  lines(sales10adj$Product_8[1:1500], type = "l", col = "pink") +
                  lines(sales10adj$Product_9[1:1500], type = "l", col = "black") +
                  lines(sales10adj$Product_10[1:1500], type = "l", col = "yellow") + 
                  abline(h = 100)

```


In order to determine the threshold for the extreme values, we will take the 98% quantile of the distribution of the sales of each product, as we could find very different distributions and taking only one threshold overall would not give us any important information. 

#### Chi-plot by couple of products {-}

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.asp=2}

par(mfrow=c(5,2))
for(i in 1:10){
 plot(sales10adj[1:1500,i], 
      type = "l",
      main=names(sales10adj[i]),
      xlab = "Day", 
      ylab = "Daily sales (unit)") +
  abline(h=quantile(sales10adj[,i], 0.99), col="red")
}

```

As we can see, we have some observations in each product that are above the line. We are particularly interested in seeing if there is a repetition in the observations that are above the lines that happen around the same time for different products.

#### Repitition Day {-}

Now, we will determine the number of the observation in which we pass the line for each products and we try to see if there are some days that are close enough or repetead. 

```{r, fig.width=20, echo=FALSE}

VAR1 <- ifelse(sales10adj[,1] < quantile(sales10adj[,1], 0.99), 0, 1)
date.p1 <- sales10adj %>% 
                mutate(VAR = VAR1, Date1 = (1:n())) %>%  
                filter(VAR == 1) %>% 
                dplyr::select(Date1)

VAR2 <- ifelse(sales10adj[,2] < quantile(sales10adj[,2], 0.99), 0, 1)
date.p2 <- sales10adj %>% 
                mutate(VAR = VAR2, Date2 = (1:n())) %>%  
                filter(VAR == 1) %>% 
                dplyr::select(Date2)

VAR3 <- ifelse(sales10adj[,3] < quantile(sales10adj[,3], 0.99), 0, 1)
date.p3 <- sales10adj %>% 
              mutate(VAR = VAR3, Date3 = (1:n())) %>%  
              filter(VAR == 1) %>% 
              dplyr::select(Date3)

VAR4 <- ifelse(sales10adj[,4] < quantile(sales10adj[,4], 0.99), 0, 1)
date.p4 <- sales10adj %>% 
              mutate(VAR = VAR4, Date4 = (1:n())) %>%  
              filter(VAR == 1) %>% 
              dplyr::select(Date4)

VAR5 <- ifelse(sales10adj[,5] < quantile(sales10adj[,5], 0.99), 0, 1)
date.p5 <- sales10adj %>% 
            mutate(VAR = VAR5, Date5 = (1:n())) %>%  
            filter(VAR == 1) %>% 
            dplyr::select(Date5)

VAR6 <- ifelse(sales10adj[,6] < quantile(sales10adj[,6], 0.99), 0, 1)
date.p6 <- sales10adj %>% 
            mutate(VAR = VAR6, Date6 = (1:n())) %>%  
            filter(VAR == 1) %>% 
            dplyr::select(Date6)

VAR7 <- ifelse(sales10adj[,7] < quantile(sales10adj[,7], 0.99), 0, 1)
date.p7 <- sales10adj %>% 
              mutate(VAR = VAR7, Date7 = (1:n())) %>%  
              filter(VAR == 1) %>% 
              dplyr::select(Date7)

VAR8 <- ifelse(sales10adj[,8] < quantile(sales10adj[,8], 0.99), 0, 1)
date.p8 <- sales10adj %>% 
              mutate(VAR = VAR8, Date8 = (1:n())) %>%  
              filter(VAR == 1) %>% 
              dplyr::select(Date8)

VAR9 <- ifelse(sales10adj[,9] < quantile(sales10adj[,9], 0.99), 0, 1)
date.p9 <- sales10adj %>% 
            mutate(VAR = VAR9, Date9 = (1:n())) %>%  
            filter(VAR == 1) %>% 
            dplyr::select(Date9)

VAR10 <- ifelse(sales10adj[,10] < quantile(sales10adj[,10], 0.99), 0, 1)
date.p10 <- sales10adj %>% 
              mutate(VAR = VAR10, Date10 = (1:n())) %>%  
              filter(VAR == 1) %>% 
              dplyr::select(Date10)

dates <- cbind(date.p1, date.p2, date.p3, date.p4, 
               date.p5, date.p6, date.p7, date.p8, 
               date.p9, date.p10)

dates<- dates%>%
            mutate(rday = case_when(Date1 == Date2 ~ 1, 
                                    Date1 == Date3 ~ 1 ,
                                    Date1 == Date4 ~ 1 ,
                                    Date1 == Date5 ~ 1 ,
                                    Date1 == Date6 ~ 1 ,
                                    Date1 == Date7 ~ 1 ,
                                    Date1 == Date8 ~ 1 ,
                                    Date1 == Date9 ~ 1 ,
                                    Date1 == Date10 ~ 1 ,
                                    Date2 == Date3 ~ 1 ,
                                    Date2 == Date4 ~ 1 ,
                                    Date2 == Date5 ~ 1 ,
                                    Date2 == Date6 ~ 1 ,
                                    Date2 == Date7 ~ 1 ,
                                    Date2 == Date8 ~ 1 ,
                                    Date2 == Date9 ~ 1 ,
                                    Date2 == Date10 ~ 1,
                                    Date3 == Date4 ~ 1 ,
                                    Date3 == Date5 ~ 1 ,
                                    Date3 == Date6 ~ 1 ,
                                    Date3 == Date7 ~ 1 ,
                                    Date3 == Date8 ~ 1 ,
                                    Date3 == Date9 ~ 1 ,
                                    Date3 == Date10 ~ 1 ,
                                    Date4 == Date5 ~ 1 ,
                                    Date4 == Date6 ~ 1 ,
                                    Date4 == Date7 ~ 1 ,
                                    Date4 == Date8 ~ 1 ,
                                    Date4 == Date9 ~ 1 ,
                                    Date4 == Date10 ~ 1 ,
                                    Date5 == Date6 ~ 1 ,
                                    Date5 == Date6 ~ 1 ,
                                    Date5 == Date7 ~ 1 ,
                                    Date5 == Date8 ~ 1 ,
                                    Date5 == Date9 ~ 1 ,
                                    Date5 == Date10 ~ 1 ,
                                    Date6 == Date7 ~ 1 ,
                                    Date6 == Date8 ~ 1 ,
                                    Date6 == Date9 ~ 1 ,
                                    Date6 == Date10 ~ 1,
                                    Date7 == Date8 ~ 1 ,
                                    Date7 == Date9 ~ 1 ,
                                    Date7 == Date10 ~ 1 ,
                                    Date8 == Date9 ~ 1 ,
                                    Date8 == Date10 ~ 1 ,
                                    Date9 == Date10 ~ 1 ,
                                     TRUE ~ 0))

names(dates) <- c("Product 1","Product 2", "Product 3",
                      "Product 4","Product 5", "Product 6",
                      "Product 7","Product 8", "Product 9",
                      "Product 10")


kableExtra::kable(dates[,1:10], 
                caption = "Check if there are any repetition above the lines that happen around the same time between the products")%>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  row_spec(c(2,3,4,5,6,7,8,10,11,20), background = "#D7261E", color = "white")

```

We can see that especially for the dates of **products 2 and 3** we have some repetitions of the same values, also for products 8 and 7 it looks like there is some similitudes or closeness of the values. In addition, the rows highlighted in red share the same day in which the extreme observation occurred.

#### Chi-plot {-}

The other method we can use is to create the chi plot for all the couples of products to assess the distributions of their tails and see if there is any correlation. 

```{r , echo=FALSE, fig.asp=2, warning=FALSE, message=FALSE}

tvec <- c("p2", "p3", "p4", "p5", "p6", "p7", "p8", "p9", "p10")

par(mfrow=c(5,2))
for(i in 1:9){
  n <- sales10adj[, c(i,(i+1))]
  evd::chiplot(n, which = 1, 
               main1 = substitute(paste('p1', a), 
              list(a = tvec[i]))) }

par(mfrow=c(4,2))
for(i in 1:8){
  n <- sales10adj[, c(i,(i+2))]
  evd::chiplot(n, which = 1, main1 = substitute(paste('p2', a), list(a = tvec[i+1])))
}

par(mfrow=c(4,2))
for(i in 1:7){
  n <- sales10adj[, c(i,(i+3))]
  evd::chiplot(n, which = 1, main1 = substitute(paste('p3', a), list(a = tvec[i+2])))
}


```


```{r , echo=FALSE, fig.asp=1.5, warning=FALSE, message=FALSE}

par(mfrow=c(3,2))
for(i in 1:6){
  n <- sales10adj[, c(i,(i+4))]
  evd::chiplot(n, which = 1, main1 = substitute(paste('p4', a), list(a = tvec[i+3])))
}

par(mfrow=c(3,2))
for(i in 1:5){
  n <- sales10adj[, c(i,(i+5))]
  evd::chiplot(n, which = 1, main1 = substitute(paste('p5', a), list(a = tvec[i+4])))
}

par(mfrow=c(4,1))
for(i in 1:4){
  n <- sales10adj[, c(i,(i+6))]
  evd::chiplot(n, which = 1, main1 = substitute(paste('p6', a), list(a = tvec[i+5])))
  }

par(mfrow=c(3,1))
for(i in 1:3){
  n <- sales10adj[, c(i,(i+7))]
  evd::chiplot(n, which = 1, main1 = substitute(paste('p7', a), list(a = tvec[i+6])))
  }

```



```{r , echo=FALSE, fig.asp=1, warning=FALSE, message=FALSE}

par(mfrow=c(2,1))
for(i in 1:2){
  n <- sales10adj[, c(i,(i+8))]
  evd::chiplot(n, which = 1, main1 = substitute(paste('p8', a), list(a = tvec[i+7])))
}

for(i in 1:1){
  n <- sales10adj[, c(i,(i+9))]
  evd::chiplot(n, which = 1, main1 = substitute(paste('p9', a), list(a = tvec[i+8])))
}

```

### {.toc-ignore}

**Key takeaway**

We can see that in the highest quantiles we have a straight line for the majority of products, only for a couple of them (like p1p3 and p5p6), there seems to be a line that is not centered around zero, but to a higher value, meaning that there could be some correlation among them, which could be a problem.


### Matematical method

While for the mathematical method, we will construct a matrix containing the chi bar values for all the couples of products. 

```{r Kable of tail dependance, echo=FALSE, warning=FALSE, message=FALSE}

# Computing the matrix of chibar values (evaluating the tail dependances among products)

P1<- list(TaildepP1P2 = taildep(sales10adj$Product_1, sales10adj$Product_2, 
                                0.95,type = c("chibar"), na.rm = FALSE), 
         TaildepP1P3 = taildep(sales10adj$Product_1, sales10adj$Product_3, 
                               0.95,type = c("chibar"), na.rm = FALSE),
         TaildepP1P4 = taildep(sales10adj$Product_1, sales10adj$Product_4, 
                               0.95,type = c("chibar"), na.rm = FALSE),
         TaildepP1P5 = taildep(sales10adj$Product_1, sales10adj$Product_5, 
                               0.95,type = c("chibar"), na.rm = FALSE),
         TaildepP1P6 = taildep(sales10adj$Product_1, sales10adj$Product_6, 
                               0.95,type = c("chibar"), na.rm = FALSE),
         TaildepP1P7 = taildep(sales10adj$Product_1, sales10adj$Product_7, 
                               0.95,type = c("chibar"), na.rm = FALSE),
         TaildepP1P8 = taildep(sales10adj$Product_1, sales10adj$Product_8, 
                               0.95,type = c("chibar"), na.rm = FALSE),
         TaildepP1P9 = taildep(sales10adj$Product_1, sales10adj$Product_9, 
                               0.95,type = c("chibar"), na.rm = FALSE),
         TaildepP1P10 = taildep(sales10adj$Product_1, sales10adj$Product_10,
                                0.95,type = c("chibar"), na.rm = FALSE))

P2<- list(TaildepP2P3 = 0,
          TaildepP2P3 =taildep(sales10adj$Product_2,sales10adj$Product_3,
                               0.95,type = c("chibar"), na.rm = FALSE),
         TaildepP2P4 =taildep(sales10adj$Product_2,sales10adj$Product_4,
                              0.95,type = c("chibar"), na.rm = FALSE),
         TaildepP2P5 =taildep(sales10adj$Product_2,sales10adj$Product_5,
                              0.95,type = c("chibar"), na.rm = FALSE),
         TaildepP2P6 =taildep(sales10adj$Product_2,sales10adj$Product_6,
                              0.95,type = c("chibar"), na.rm = FALSE),
         TaildepP2P7 =taildep(sales10adj$Product_2,sales10adj$Product_7,
                              0.95,type = c("chibar"), na.rm = FALSE),
         TaildepP2P8 =taildep(sales10adj$Product_2,sales10adj$Product_8,
                              0.95,type = c("chibar"), na.rm = FALSE),
         TaildepP2P9 =taildep(sales10adj$Product_2,sales10adj$Product_9,
                              0.95,type = c("chibar"), na.rm = FALSE),
         TaildepP2P10 =taildep(sales10adj$Product_2,sales10adj$Product_10,
                               0.95,type = c("chibar"), na.rm = FALSE))

P3 <-list(Taildep = 0,
          Taildep = 0,
          TaildepP4P4 =taildep(sales10adj$Product_3,sales10adj$Product_4,
                               0.95,type = c("chibar"), na.rm = FALSE),
            TaildepP4P5 =taildep(sales10adj$Product_3,sales10adj$Product_5,
                                 0.95,type = c("chibar"), na.rm = FALSE),
            TaildepP4P6 =taildep(sales10adj$Product_3,sales10adj$Product_6,
                                 0.95,type = c("chibar"), na.rm = FALSE),
            TaildepP4P7 =taildep(sales10adj$Product_3,sales10adj$Product_7,
                                 0.95,type = c("chibar"), na.rm = FALSE),
            TaildepP4P8 =taildep(sales10adj$Product_3,sales10adj$Product_8,
                                 0.95,type = c("chibar"), na.rm = FALSE),
            TaildepP4P9 =taildep(sales10adj$Product_3,sales10adj$Product_9,
                                 0.95,type = c("chibar"), na.rm = FALSE),
            TaildepP4P10 =taildep(sales10adj$Product_3,sales10adj$Product_10,
                                  0.95,type = c("chibar"), na.rm = FALSE))

P4 <-list(Taildep = 0,
          Taildep = 0,
          Taildep = 0,
          TaildepP4P5 =taildep(sales10adj$Product_4,sales10adj$Product_5,
                               0.95,type = c("chibar"), na.rm = FALSE),
          TaildepP4P6 =taildep(sales10adj$Product_4,sales10adj$Product_6,
                               0.95,type = c("chibar"), na.rm = FALSE),
          TaildepP4P7 =taildep(sales10adj$Product_4,sales10adj$Product_7,
                               0.95,type = c("chibar"), na.rm = FALSE),
          TaildepP4P8 =taildep(sales10adj$Product_4,sales10adj$Product_8,
                               0.95,type = c("chibar"), na.rm = FALSE),
          TaildepP4P9 =taildep(sales10adj$Product_4,sales10adj$Product_9,
                               0.95,type = c("chibar"), na.rm = FALSE),
          TaildepP4P10 =taildep(sales10adj$Product_4,sales10adj$Product_10,
                                0.95,type = c("chibar"), na.rm = FALSE))

P5 <-list(Taildep = 0,
          Taildep = 0,
          Taildep = 0,
          Taildep=  0,
          TaildepP4P6 =taildep(sales10adj$Product_5,sales10adj$Product_6,
                               0.95,type = c("chibar"), na.rm = FALSE),
          TaildepP4P7 =taildep(sales10adj$Product_5,sales10adj$Product_7,
                               0.95,type = c("chibar"), na.rm = FALSE),
          TaildepP4P8 =taildep(sales10adj$Product_5,sales10adj$Product_8,
                               0.95,type = c("chibar"), na.rm = FALSE),
          TaildepP4P9 =taildep(sales10adj$Product_5,sales10adj$Product_9,
                               0.95,type = c("chibar"), na.rm = FALSE),
          TaildepP4P10 =taildep(sales10adj$Product_5,sales10adj$Product_10,
                                0.95,type = c("chibar"), na.rm = FALSE))

P6 <-list(Taildep = 0,
          Taildep = 0,
          Taildep = 0,
          Taildep=  0,
          Taildep=  0,
          TaildepP4P7 =taildep(sales10adj$Product_6,sales10adj$Product_7,
                               0.95,type = c("chibar"), na.rm = FALSE),
          TaildepP4P8 =taildep(sales10adj$Product_6,sales10adj$Product_8,
                               0.95,type = c("chibar"), na.rm = FALSE),
          TaildepP4P9 =taildep(sales10adj$Product_6,sales10adj$Product_9,
                               0.95,type = c("chibar"), na.rm = FALSE),
          TaildepP4P10 =taildep(sales10adj$Product_6,sales10adj$Product_10,
                                0.95,type = c("chibar"), na.rm = FALSE))

P7 <-list(Taildep = 0,
          Taildep = 0,
          Taildep = 0,
          Taildep=  0,
          Taildep=  0,
          Taildep=  0,
          TaildepP4P8 =taildep(sales10adj$Product_7,sales10adj$Product_8,
                               0.95,type = c("chibar"), na.rm = FALSE),
          TaildepP4P9 =taildep(sales10adj$Product_7,sales10adj$Product_9,
                               0.95,type = c("chibar"), na.rm = FALSE),
          TaildepP4P10 =taildep(sales10adj$Product_7,sales10adj$Product_10,
                                0.95,type = c("chibar"), na.rm = FALSE))

P8 <-list(Taildep = 0,
          Taildep = 0,
          Taildep = 0,
          Taildep=  0,
          Taildep=  0,          
          Taildep=  0,
          Taildep=  0,
          TaildepP4P9 =taildep(sales10adj$Product_8,sales10adj$Product_9,
                               0.95,type = c("chibar"), na.rm = FALSE),
          TaildepP4P10 =taildep(sales10adj$Product_8,sales10adj$Product_10,
                                0.95,type = c("chibar"), na.rm = FALSE))

P9 <-list(Taildep = 0,
          Taildep = 0,
          Taildep = 0,
          Taildep=  0,
          Taildep=  0,          
          Taildep=  0,
          Taildep=  0,
          Taildep=  0,
          TaildepP4P10 =taildep(sales10adj$Product_9,sales10adj$Product_10,
                                0.95,type = c("chibar"), na.rm = FALSE))

pp<- as.data.frame(rbind(unlist(P1),unlist(P2),unlist(P3),
                         unlist(P4),unlist(P5),unlist(P6),
                         unlist(P7),unlist(P8),unlist(P9)))
pp<-round(pp,3)

colnames(pp) <- c("p2", "p3", "p4", "p5", "p6", "p7", "p8", "p9", "p10") 
rownames(pp) <- c("p1", "p2", "p3", "p4", "p5", "p6", "p7", "p8", "p9") 

pp[1:9, 1:9] %>%
  rownames_to_column("pp") %>% 
  mutate(p2 = cell_spec(p2, color = ifelse(abs(p2) > 0.4, "green", "#D7261E"))) %>% 
  mutate(p3 = cell_spec(p3, color = ifelse(abs(p3) > 0.4, "green", "#D7261E"))) %>% 
  mutate(p4 = cell_spec(p4, color = ifelse(abs(p4) > 0.4, "green", "#D7261E"))) %>% 
  mutate(p5 = cell_spec(p5, color = ifelse(abs(p5) > 0.4, "green", "#D7261E"))) %>% 
  mutate(p6 = cell_spec(p6, color = ifelse(abs(p6) > 0.4, "green", "#D7261E"))) %>% 
  mutate(p7 = cell_spec(p7, color = ifelse(abs(p7) > 0.4, "green", "#D7261E"))) %>% 
  mutate(p8 = cell_spec(p8, color = ifelse(abs(p8) > 0.4, "green", "#D7261E"))) %>% 
  mutate(p9 = cell_spec(p9, color = ifelse(abs(p9) > 0.4, "green", "#D7261E"))) %>% 
  mutate(p10 = cell_spec(p10, color = ifelse(abs(p10) > 0.4, "green", "#D7261E"))) %>% 
  kableExtra::kable(booktabs = T, linesep = "", escape=FALSE, caption = "Chi bar Matrix") %>% 
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

```

We can see that the highest values appear to be for *products 2* with *product 3 and 4*, *product 1 with product 5 and 6*, product 3 with 4, *products 5 with 6.* This is very much in line with we have found in the chi plot that we created above. 

*Correlations greater than 0.4 are highlighted in green.*

### Aggregation method {.tabset .tabset-fade .tabset-pills}

To investigate a possible clustering of the different products, we will create create a graph with the mean and variance for each product. 

#### Distance: Mean and variance

```{r, fig.width=10, include=FALSE}

#### For all the data #####################

var.sales10adj <- diag(var(sales10adj))    # compute variance
mean.sales10adj <-apply(sales10adj, 2, mean)   # compute mean

#### For extreme values#####################

sales1 <- sales10adj %>% mutate(VAR = VAR1) %>%  filter(VAR == 1) %>% dplyr::select(Product_1)
sales2 <- sales10adj %>% mutate(VAR = VAR2) %>%  filter(VAR == 1) %>% dplyr::select(Product_2)
sales3 <- sales10adj %>% mutate(VAR = VAR3) %>%  filter(VAR == 1) %>% dplyr::select(Product_3)
sales4 <- sales10adj %>% mutate(VAR = VAR4) %>%  filter(VAR == 1) %>% dplyr::select(Product_4)
sales5 <- sales10adj %>% mutate(VAR = VAR5) %>%  filter(VAR == 1) %>% dplyr::select(Product_5)
sales6 <- sales10adj %>% mutate(VAR = VAR6) %>%  filter(VAR == 1) %>% dplyr::select(Product_6)
sales7 <- sales10adj %>% mutate(VAR = VAR7) %>%  filter(VAR == 1) %>% dplyr::select(Product_7)
sales8 <- sales10adj %>% mutate(VAR = VAR8) %>%  filter(VAR == 1) %>% dplyr::select(Product_8)
sales9 <- sales10adj %>% mutate(VAR = VAR9) %>%  filter(VAR == 1) %>% dplyr::select(Product_9)
sales10 <- sales10adj %>% mutate(VAR = VAR10) %>%  filter(VAR == 1) %>% dplyr::select(Product_10)

saleshigh <- cbind(sales1, sales2, sales3, sales4, sales5, 
                   sales6, sales7, sales8, sales9, sales10)

var.saleshigh <- diag(var(saleshigh))    # compute variance
mean.saleshigh <-apply(saleshigh, 2, mean)   # compute mean

```

```{r, echo=FALSE, fig.width=10, warning=FALSE, message=FALSE, fig.asp=1.2}

par(mfrow=c(2,1))
plot(var.sales10adj ~ mean.sales10adj,       #plot with all the data
     main = "Similarity between variables all dataset", 
     xlab = "Mean value per product", 
     ylab = "Variance per product",
     xlim = c(-0.005,0.007),
     ylim = c(0, 1700))
     text(var.sales10adj ~ mean.sales10adj, labels=c("p1", "p2", "p3", "p4", 
                                                     "p5", "p6", "p7", "p8", 
                                                     "p9", "p10"),
          data=sales10adj, pos=3, offset=0.3, cex=1, font=1)
plot(var.saleshigh ~ mean.saleshigh,         #plot for extreme values
     main = "Similarity between variables extreme values", 
     xlab = "Mean value per product", 
     ylab = "Variance per product")
     text(var.saleshigh ~ mean.saleshigh, 
          labels=c("p1", "p2", "p3", "p4", "p5", 
                   "p6", "p7", "p8", "p9", "p10"),
          data=saleshigh, pos=3, offset=0.3, cex=1, font=1)   
     
```

#### Defining k

```{r, echo=FALSE}

#Data Adjust transposed and scale
sales10adj_1<-as.data.frame(rbind(sales10adj$Product_1, sales10adj$Product_2,
                                          sales10adj$Product_3, sales10adj$Product_4,
                                          sales10adj$Product_5, sales10adj$Product_6,
                                          sales10adj$Product_7, sales10adj$Product_8,
                                          sales10adj$Product_9, sales10adj$Product_10))
#scale the data
sales10adj_1<-scale(sales10adj_1)

#Computing distance
distance_1<- factoextra::get_dist(sales10adj_1,
                                method = "kendall")
#Defining the k
factoextra::fviz_nbclust(sales10adj_1,
                         kmeans,
                         method = "wss",       #for total within sum of square--> "wss"
                         k.max = 8,
                         diss = distance_1,      #kendall distance
                         nstart = 50)    + 
       geom_vline(xintercept = 3, color="red")+
  labs(subtitle = "The elbow begins to stabilize from cluster 3")+
  theme(plot.subtitle=element_text(face="italic"))+
  theme(plot.title=element_text(size=14, hjust=0.5, face="bold"))

```


#### Correlation: Kendall's tau

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.asp=0.6}

#create the cluster with all the data
cluster_k3_1 <- kmeans(sales10adj_1, 3 , nstart = 25)

#Correlation of the total dataset
CoR_1<- factoextra::fviz_dist(distance_1) #correlation
#plot of the total data set
g1<-factoextra::fviz_cluster(cluster_k3_1, data = sales10adj_1, show.clust.cent = TRUE,
             star.plot = TRUE, repel = TRUE) +
      ggplot2::labs(title = "Clustering our 10 products by their Kendall's tau distance values)") +
      ggplot2::theme_bw() +
    ggplot2::theme(legend.position = "bottom")+
    ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5))
  
######now for the extreme values###################

####create the data frame
q95<- as.data.frame(apply(sales10adj, 2, quantile, 0.95))  #quantile 95
Product_1<- sales10adj%>%
                  dplyr::filter(sales10adj$Product_1 >= q95[1,1])%>%
                  dplyr::select(Product_1)
Product_2<- sales10adj%>%
                  filter(sales10adj$Product_2 >= q95[2,1])%>%
                  dplyr::select(Product_2)
Product_3<- sales10adj%>%
                  filter(sales10adj$Product_3 >= q95[3,1])%>%
                  dplyr::select(Product_3)
Product_4<- sales10adj%>%
                  filter(sales10adj$Product_4 >= q95[4,1])%>%
                  dplyr::select(Product_4)
Product_5<- sales10adj%>%
                  filter(sales10adj$Product_5 >= q95[5,1])%>%
                  dplyr::select(Product_5)
Product_6<- sales10adj%>%
                  filter(sales10adj$Product_6 >= q95[6,1])%>%
                  dplyr::select(Product_6)
Product_7<- sales10adj%>%
                  filter(sales10adj$Product_7 >= q95[7,1])%>%
                  dplyr::select(Product_7)
Product_8<- sales10adj%>%
                  filter(sales10adj$Product_8 >= q95[8,1])%>%
                  dplyr::select(Product_8)
Product_9<- sales10adj%>%
                  filter(sales10adj$Product_9 >= q95[9,1])%>%
                  dplyr::select(Product_9)
Product_10<- sales10adj%>%
                  filter(sales10adj$Product_10 >= q95[10,1])%>%
                  dplyr::select(Product_10)

sales10adj_2<- data.frame(Product_1, Product_2, Product_3,
                          Product_4, Product_5, Product_6,
                          Product_7, Product_8, Product_9,
                          Product_10)
sales10adj_2<-as.data.frame(rbind(sales10adj_2$Product_1, sales10adj_2$Product_2,
                                          sales10adj_2$Product_3, sales10adj_2$Product_4,
                                          sales10adj_2$Product_5, sales10adj_2$Product_6,
                                          sales10adj_2$Product_7, sales10adj_2$Product_8,
                                          sales10adj_2$Product_9, sales10adj_2$Product_10))
#scale
sales10adj_2<-scale(sales10adj_2)

##########cluster extreme
distance_2<- factoextra::get_dist(sales10adj_2, 
                                method = "kendall")

CoR_2<-factoextra::fviz_dist(distance_2) #correlation

cluster_k3_2 <- kmeans(sales10adj_2, 3 , nstart = 25)

g2<- factoextra::fviz_cluster(cluster_k3_2, data = sales10adj_2, show.clust.cent = TRUE,
             star.plot = TRUE, repel = TRUE) +
      ggplot2::labs(title = "Clustering of the extreme values of our 10 products") +
      ggplot2::theme_bw() +
      ggplot2::theme(legend.position = "bottom")+
      ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5))

gridExtra::grid.arrange(CoR_1, CoR_2, nrow=1, 
                        top = "Correlation all data VS extreme values")
```

#### Cluster with distance: Kendall's tau

```{r , warning=FALSE, message=FALSE, echo=FALSE, fig.asp=1.5}

gridExtra::grid.arrange(g1, g2, ncol = 1)

```


### {.toc-ignore}

**Key takeaway**

What is interesting to notice, is that the clustering are the same if we consider the whole distribution and only the most extreme values. Netherless, the correlation are stronger in the extreme values, thus we confirm that an independent analysis could potentialy lose important information.

In both cases, we can see that there is some similarity in some products, and some differences in others, especially product 1 appreas to be very different from the others, while products 9, 10 7, 8, 4, 3 and 2 appear very similar (the cluster in green). However, the number of clusters that should be used is not very clear, as the decrease of total within sum of sqaure is quite linear. 

## Exercise C

> In view of your answer to (b), would you apply a Gaussian copula model to these data?

As well as the **EVT** the Gaussian model does not caputre all the information for the extreme values. Keeping this concept for multivariate extreme value theory we can see that in this case the corelation is higher, thus we cannot apply this model to fit the extreme value of the products' distribution. 

As we could find some depence in the distribution of the tails of our products' demand, we believe that a gaussian distribution would not be optimal, as it is not able to consider the depences in the tails of the distribution.

## Exercise D

> Fit a multivariate model to the adjusted sales data and estimate the 95% Demand at Risk for the sum:

$$S = X^{(1)} + X^{(2)} + ... + X^{(10)}$$

You can do so with the following steps: 

```{r, warning=FALSE, message=FALSE}

# These functions can be used to transform data to uniform scale,
# when the data are analysed separately below and above a threshold
# e.g. applying the GPD to values above a high threshold and treating
#      those below the threshold with the empirical CDF.

transform_to_uniform <- function(y,
                                 cdf,
                                 threshold,
                                 ...) {
  # ...: arguments to `cdf`
  
  
  above <- y > threshold
  # proportion of exceedances
  p <- mean(above)
  # under the threshold, use the empirical CDF
  # see ?ecdf
  ecdf_below <- ecdf(y[!above])
  empirical <- ecdf_below(y[!above])
  # above, apply cdf
  theoretical <- cdf(y[above], mean.sales10adj)
  
  transformed <- numeric(length = length(y))
  # "Glue together" the empirical and theoretical parts by rescaling them:
  
  #        empirical          e.g. N(0,1)
  # [-----------------------|-------------]
  # 0                      1-p            1
  
  transformed[!above] <- (1 - p) * empirical    # (1)
  transformed[above] <- 1 - p + p * theoretical # (2)
  
  return(list(
    transformed = transformed,
    ecdf = ecdf_below,
    prop = p
  ))
}

inverse_transform <- function(u, ecdf, quantile_function, p, ...) {
  # ...: arguments to quantile_function
  # Basically, undo the transformations (1-2) above
  above <- u > 1 - p
  original_scale <- numeric(length = length(u))
  original_scale[!above] <- quantile(ecdf, u[!above] / (1 - p))
  original_scale[above] <-
    quantile_function((u[above] - (1 - p)) / p, ...)
  return(original_scale)
}

```

i. Transform the sales for each product to a uniform scale, that is, compute 
$$F_i(X^{(i)})$$
for each X in 1, ..., n being each of the product sales and F being the estimated Extreme Value distributions.


```{r, warning=FALSE, echo=FALSE, message=FALSE}

y1 <- sales10adj$Product_1
th1 <- qpois(0.95, abs(mean(sales10adj$Product_1)))
unif1 <- transform_to_uniform(y1, ppois, th1)

y2 <- sales10adj$Product_2
th2 <- qpois(0.95, abs(mean.sales10adj[2]))
unif2 <- transform_to_uniform(y2, ppois, th2)

y3 <- sales10adj$Product_3
th3 <- qpois(0.95, abs(mean.sales10adj[3]))
unif3 <- transform_to_uniform(y3, ppois, th3)

y4 <- sales10adj$Product_4
th4 <- qpois(0.95, mean.sales10adj[4])
unif4 <- transform_to_uniform(y4, ppois, th4)

y5 <- sales10adj$Product_5
th5 <- qpois(0.95, mean.sales10adj[5])
unif5 <- transform_to_uniform(y5, ppois, th5)

y6 <- sales10adj$Product_6
th6 <- qpois(0.95, abs(mean.sales10adj[6]))
unif6 <- transform_to_uniform(y6, ppois, th6)

y7 <- sales10adj$Product_7
th7 <- qpois(0.95, abs(mean.sales10adj[7]))
unif7 <- transform_to_uniform(y7, ppois, th7)

y8 <- sales10adj$Product_8
th8 <- qpois(0.95, abs(mean.sales10adj[8]))
unif8 <- transform_to_uniform(y8, ppois, th8)

y9 <- sales10adj$Product_9
th9 <- qpois(0.95, abs(mean.sales10adj[9]))
unif9 <- transform_to_uniform(y9, ppois, th9)

y10 <- sales10adj$Product_10
th10 <- qpois(0.95, abs(mean.sales10adj[10]))
unif10 <- transform_to_uniform(y10, ppois, th10)

```


ii. Fit a multivariate copula of your choice using the `copula::fitCopula` function.

```{r, warning=FALSE, message=FALSE}
library(copula)

unif <- cbind(unif1$transformed, unif2$transformed, unif3$transformed, unif4$transformed, unif5$transformed, unif6$transformed, unif7$transformed, unif8$transformed, unif9$transformed, unif10$transformed)

gum.fit <- gumbelCopula(5, dim=10)
clay.fit <- claytonCopula(4, dim=10)

fit.gum <- fitCopula(gum.fit, method = "itau", data=unif)
summary(fit.gum)

```

iii. Simulate from your fitted copula and transform the values back to their original scales (i.e. undo the transformation in i.)

```{r, fig.asp=2, warning=FALSE, echo=FALSE, message=FALSE}
library(copula)
sim.gum <- rCopula(copula = gumbelCopula(1.04, 10), n = 1000)

inverse_transform <- function(u, ecdf, quantile_function, p, ...) {
  # ...: arguments to quantile_function
  # Basically, undo the transformations (1-2) above
  above <- u > 1 - p
  original_scale <- numeric(length = length(u))
  original_scale[!above] <- quantile(ecdf, u[!above] / (1 - p))
  original_scale[above] <-
    quantile_function((u[above] - (1 - p)) / p, mean.sales10adj)
  return(original_scale)
}

rev1 <-  with(unif1, inverse_transform(sim.gum[,1], ecdf, qnorm, p = prop))
rev2 <-  with(unif2, inverse_transform(sim.gum[,2], ecdf, qnorm, p = prop))
rev3 <-  with(unif3, inverse_transform(sim.gum[,3], ecdf, qnorm, p = prop))
rev4 <-  with(unif4, inverse_transform(sim.gum[,4], ecdf, qnorm, p = prop))
rev5 <-  with(unif5, inverse_transform(sim.gum[,5], ecdf, qnorm, p = prop))
rev6 <-  with(unif6, inverse_transform(sim.gum[,6], ecdf, qnorm, p = prop))
rev7 <-  with(unif7, inverse_transform(sim.gum[,7], ecdf, qnorm, p = prop))
rev8 <-  with(unif8, inverse_transform(sim.gum[,8], ecdf, qnorm, p = prop))
rev9 <-  with(unif9, inverse_transform(sim.gum[,9], ecdf, qnorm, p = prop))
rev10 <-  with(unif10, inverse_transform(sim.gum[,10], ecdf, qnorm, p = prop))

par(mfrow=c(5,2))
plot(sales10adj[1:1000,1], rev1)
plot(sales10adj[1:1000,2], rev2)
plot(sales10adj[1:1000,3], rev3)
plot(sales10adj[1:1000,4], rev4)
plot(sales10adj[1:1000,5], rev5)
plot(sales10adj[1:1000,6], rev6)
plot(sales10adj[1:1000,7], rev7)
plot(sales10adj[1:1000,8], rev8)
plot(sales10adj[1:1000,9], rev9)
plot(sales10adj[1:1000,10], rev10)


```

iv. Compute the simulated values of S and their 95% Value at Risk.

```{r, fig.asp=2, warning=FALSE, echo=FALSE, message=FALSE}
par(mfrow=c(5,2))
plot(rev1, type = "l")
plot(rev2, type = "l")
plot(rev3, type = "l")
plot(rev4, type = "l")
plot(rev5, type = "l")
plot(rev6, type = "l")
plot(rev7, type = "l")
plot(rev8, type = "l")
plot(rev9, type = "l")
plot(rev10, type = "l")
  
q95p1 <- quantile(rev1, 0.95)
VAR1 <- as.data.frame(rev1) %>%
                  dplyr::filter(rev1 >= q95p1)

q95p2 <- quantile(rev2, 0.95)
VAR2 <- as.data.frame(rev2) %>%
                  dplyr::filter(rev2 >= q95p2)

q95p3 <- quantile(rev3, 0.95)
VAR3 <- as.data.frame(rev3) %>%
                  dplyr::filter(rev3 >= q95p3)

q95p4 <- quantile(rev4, 0.95)
VAR4 <- as.data.frame(rev4) %>%
                  dplyr::filter(rev4 >= q95p4)

q95p5 <- quantile(rev5, 0.95)
VAR5 <- as.data.frame(rev5) %>%
                  dplyr::filter(rev5 >= q95p5)

q95p6 <- quantile(rev6, 0.95)
VAR6 <- as.data.frame(rev6) %>%
                  dplyr::filter(rev6 >= q95p6)

q95p7 <- quantile(rev7, 0.95)
VAR7 <- as.data.frame(rev7) %>%
                  dplyr::filter(rev7 >= q95p7)

q95p8 <- quantile(rev8, 0.95)
VAR8 <- as.data.frame(rev8) %>%
                  dplyr::filter(rev8 >= q95p8)

q95p9 <- quantile(rev9, 0.95)
VAR9 <- as.data.frame(rev9) %>%
                  dplyr::filter(rev9 >= q95p9)

q95p10 <- quantile(rev10, 0.95)
VAR10 <- as.data.frame(rev10) %>%
                  dplyr::filter(rev10 >= q95p10)
str(VAR10)

VAR <- cbind(VAR1, VAR2, VAR3, VAR4, VAR5, VAR6, VAR7, VAR8, VAR9, VAR10) 

VAR<- round(VAR, 2)

kableExtra::kable(VAR , 
                  caption = "Extreme Values Data frame")%>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"))


```